---
title: "exercise-3"
author: "Yen"
format: html
editor: visual
---

## Preliminaries

getwd()

str(d)

names()

## Running Code

```{r}
d <- read.csv("https://raw.githubusercontent.com/yendo3877/exercise-03/refs/heads/main/data-wrangling.csv", header = TRUE)
names(d)
```

## 1-3. Create variables

```{r}
BSD <- d$Body_mass_male_mean/d$Body_mass_female_mean
sex_ratio <- d$AdultFemale/d$AdultMales
DI <- d$DayLength_km/(sqrt(d$HomeRange_km2/pi))

```

## 4-6. Plot

```{r}
library(ggplot2)
library(dplyr)
```

#### 4. Relationship between day range length (y axis) and time spent moving (x axis)

```{r}
p <- ggplot(data = d, aes(x = log(Move), y = log(DayLength_km),
    color = "black"))  
p <- p + xlab("log(Move)") + ylab("log(Daylength_km)")
p <- p + geom_point(na.rm = TRUE)
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p <- p + geom_smooth(method = "lm", fullrange = FALSE, na.rm = TRUE)
p
```

For each family

```{r}
p <- p + facet_wrap(~Family, ncol = 4)
p <- p + theme(legend.position = "none")
p
```

Results: Yes. Species that spend more times moving travel farther overall. Among 14 families, Hylobatidae and Atilidae have positive correlation. Variables should be transformed to create well distributed population.

#### 5. Relationship between day range length (y axis) and group size (x axis)

```{r}
p <- ggplot(data = d, aes(x = log(MeanGroupSize), y = log(DayLength_km),
    color = "blue"))  
p <- p + xlab("group size") + ylab("day range length")
p <- p + geom_point(na.rm = TRUE)
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p <- p + geom_smooth(method = "lm", fullrange = FALSE, na.rm = TRUE)
p
```

For each family

```{r}
p <- p + facet_wrap(~Family, ncol = 4)
p <- p + theme(legend.position = "none")
p
```

Results: Yes. Species that live in larger groups travel farther overall. Positive correlation is observed in 7 families. The most significant one belongs to Cercopithecidae. Variables should be transformed to create well distributed population.

##### 6. Relationship between canine size dimorphism (y axis) and body size dimorphism (x axis)

```{r}
p <- ggplot(data = d, aes(x = log(Canine_Dimorphism), y = log(BSD),
    color = "green"))  
p <- p + xlab("body size dimorphism ") + ylab("canine size dimorphism")
p <- p + geom_point(na.rm = TRUE)
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p <- p + geom_smooth(method = "lm", fullrange = FALSE, na.rm = TRUE)
p
```

For each family

```{r}
p <- p + facet_wrap(~Family, ncol = 4)
p <- p + theme(legend.position = "none")
p
```

Results: Yes. Taxa with greater size dimorphism also show greater canine dimorphism.

## 7. diet_strategy

```{r}
d <- mutate(d, diet_strategy = paste(case_when(Fruit >= 50 ~ "frugivore", Leaves >= 50 ~ "folivore", Fruit < 50 & Leaves < 50 ~ "omnivore", TRUE ~ NA)
                                        
                                     ))
d
```

```{r}
p <- ggplot(data = d %>% filter(diet_strategy != "NA"), aes(x = diet_strategy, y = log(MeanGroupSize)))
p <- p + geom_boxplot(na.rm = TRUE)
p <- p + geom_jitter(aes(color = "black"), width = 0.2, size = 2, alpha = 0.7) 
p <- p + theme(axis.text.x = element_text(angle = 90))
p <- p + ylab("log(MeanGroupSize)")
p
```

Results: No. Frugivore live in smaller groups than folivores.

## 8

```{r}
d %>%  
  mutate(Binomial = paste(Genus, Species, sep = " ")) %>%
  select(Binomial, Family, Brain_Size_Species_Mean, Body_mass_male_mean) %>%
  group_by(Family) %>%
  summarise(avgBS = mean(Brain_Size_Species_Mean, na.rm = TRUE),
    avgBM = mean(Body_mass_male_mean, na.rm = TRUE)) %>%
  arrange(avgBS) %>%
  print
            
```
